---
title: "Presentation clustering"
author: "Linh Phung"
date: "2023-11-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data cleaning


```{r }
data <- read.csv("top_insta_influencers_data.csv")
head(data)
```

```{r}
# Define the conversion function
convert_to_numeric <- function(str) {
  # Extract the last character
  last_char <- substr(str, nchar(str), nchar(str))
  # Initialize a vector to store the results
  result <- numeric(length(str))
  for (i in 1:length(str)) {
    # Handle the cases for "k," "m," and "b"
    if (last_char[i] == "k") {
      result[i] <- as.numeric(substr(str[i], 1, nchar(str[i]) - 1)) * 1e3
    } else if (last_char[i] == "m") {
      result[i] <- as.numeric(substr(str[i], 1, nchar(str[i]) - 1)) * 1e6
    } else if (last_char[i] == "b") {
      result[i] <- as.numeric(substr(str[i], 1, nchar(str[i]) - 1)) * 1e9
    } else {
      result[i] <- as.numeric(str[i])
    }
  }
  return(result)
}

# Apply the function to specific columns in the dataframe using mutate
cols_to_convert <- c(4, 5, 6, 8, 9)
data[, cols_to_convert] <- lapply(data[, cols_to_convert], convert_to_numeric)
```

```{r}
data$X60_day_eng_rate <- as.numeric(sub("%", "", data$X60_day_eng_rate))/100
df <- data[,-c(1,2,10)]
```

```{r}
new_df <- df[df$new_post_avg_like!=0,]
summary(new_df)
```

```{r}
final_df <- scale(new_df)
```

# K-means clustering

```{r}
#install.packages("dbscan")
library(dbscan)
```

```{r}
kNNdistplot(final_df, k = 3)
abline(h = 2.5, col = "red")
```

```{r}
db <- dbscan(final_df, eps = 2.2, minPts = 4)
db
```

```{r}
clean_data1 <- data.frame(final_df[which(db$cluster != 0), ])
set.seed(100)
km <- kmeans(clean_data1, centers = 3, nstart = 30)
```

```{r}
library(factoextra)
factoextra::fviz_nbclust(clean_data1, kmeans, method='silhouette', k.max=15)
```

```{r}
factoextra::fviz_nbclust(clean_data1, kmeans, method='wss', k.max=15)
```


```{r}
library(ggplot2)
library(factoextra)
fviz_cluster(km, data = clean_data1, centroids = TRUE, repel = FALSE)
```

# Descriptive Analysis

```{r}
new_df['stt'] <- rownames(new_df)
cluster <- data.frame(km$cluster)
cluster$stt <- rownames(cluster)
outcome <- merge(new_df, cluster, on="stt")
```

```{r}
# Calculate the average per cluster
avg_posts <- aggregate(outcome$posts, by=list(outcome$km.cluster), FUN=mean)

# Create a bar plot using ggplot2
ggplot(avg_posts, aes(x=Group.1, y=x)) +
  geom_bar(stat="identity", fill="blue") +
  xlab("Cluster") +
  ylab("Average Posts") +
  ggtitle("Average Posts per Cluster")
```


```{r}
# Calculate the average per cluster
avg_fl <- aggregate(outcome$followers, by=list(outcome$km.cluster), FUN=mean)

# Create a bar plot using ggplot2
ggplot(avg_fl, aes(x=Group.1, y=x)) +
  geom_bar(stat="identity", fill="blue") +
  xlab("Cluster") +
  ylab("Average Posts") +
  ggtitle("Average Followers per Cluster")
```
```{r}
# Calculate the average per cluster
avg_like <- aggregate(outcome$avg_likes, by=list(outcome$km.cluster), FUN=mean)

# Create a bar plot using ggplot2
ggplot(avg_like, aes(x=Group.1, y=x)) +
  geom_bar(stat="identity", fill="blue") +
  xlab("Cluster") +
  ylab("Average Posts") +
  ggtitle("Average Likes per Cluster")
```

```{r}
# Calculate the average per cluster
avg_eng <- aggregate(outcome$X60_day_eng_rate, by=list(outcome$km.cluster), FUN=mean)

# Create a bar plot using ggplot2
ggplot(avg_eng, aes(x=Group.1, y=x)) +
  geom_bar(stat="identity", fill="blue") +
  xlab("Cluster") +
  ylab("Average Posts") +
  ggtitle("Average Engagement in 60 days per Cluster")
```

# Outlier Analysis

Since outliers may come from big stars on Instagram, we should conduct analysis to understand them.

```{r}
outliers <- data.frame(data[which(db$cluster == 0), ])
outliers
```
```{r}
# mean value
mean_follower <- mean(data$followers, na.rm = TRUE)

# Create a bar plot using ggplot2
ggplot(outliers, aes(x=channel_info, y=followers)) +
  geom_bar(stat="identity", fill="blue") +
  xlab("Influencers") +
  ylab("Followers") +
  ggtitle("Followers per Influencers") +
  geom_hline(yintercept = mean_follower, color = "red", linetype = "dashed") + 
  geom_text(aes(x = 2, y = mean_follower, label = paste("Mean Followers: ", round(mean_follower, 2)), vjust = -1), color = "red")  # Add text for the red line
```

```{r}
# mean value
mean_post <- mean(data$posts)

# Create a bar plot using ggplot2
ggplot(outliers, aes(x=channel_info, y=posts)) +
  geom_bar(stat="identity", fill="blue") +
  xlab("Influencers") +
  ylab("Posts") +
  ggtitle("Posts per Influencers") +
  geom_hline(yintercept = mean_post, color = "red", linetype = "dashed") + 
  geom_text(aes(x = 1.5, y = mean_post, label = paste("Mean Posts: ", round(mean_post, 2)), vjust = -1), color = "red")  # Add text for the red line
```

Some of them (cristiano, kyliejenner, zendaya) have high or close to avg posts and high number of followers.

```{r}
# mean value
mean_like <- mean(data$total_likes)

# Create a bar plot using ggplot2
ggplot(outliers, aes(x=channel_info, y=total_likes)) +
  geom_bar(stat="identity", fill="blue") +
  xlab("Influencers") +
  ylab("Posts") +
  ggtitle("Likes per Influencers") +
  geom_hline(yintercept = mean_like, color = "red", linetype = "dashed") + 
  geom_text(aes(x = 1.7, y = mean_like, label = paste("Mean Likes: ", round(mean_like, 2)), vjust = -1), color = "red")  # Add text for the red line
```
Some of them (cristiano, kyliejenner, zendaya) have very high total likes and avg likes per post.

```{r}
# mean value
mean_avg_like <- mean(data$avg_likes)

# Create a bar plot using ggplot2
ggplot(outliers, aes(x=channel_info, y=avg_likes)) +
  geom_bar(stat="identity", fill="blue") +
  xlab("Influencers") +
  ylab("Posts") +
  ggtitle("Avg. Likes per post per Influencers") +
  geom_hline(yintercept = mean_avg_like , color = "red", linetype = "dashed") + 
  geom_text(aes(x = 2, y = mean_avg_like , label = paste("Mean Avg. Likes: ", round(mean_avg_like , 2)), vjust = -1), color = "red")  # Add text for the red line
```

```{r}
# mean value
mean_60_eng <- mean(data$X60_day_eng_rate, na.rm = TRUE)

# Create a bar plot using ggplot2
ggplot(outliers, aes(x=channel_info, y=X60_day_eng_rate)) +
  geom_bar(stat="identity", fill="blue") +
  xlab("Influencers") +
  ylab("Posts") +
  ggtitle("60 days engagement rate per Influencers") +
  geom_hline(yintercept = mean_60_eng , color = "red", linetype = "dashed") + 
  geom_text(aes(x = 2, y = mean_60_eng , label = paste("Mean 60 days engagement: ", round(mean_60_eng , 2)), vjust = -1), color = "red")  # Add text for the red line
```

Most of them are not active recently, except for zendaya


